\newpage
\begin{thebibliography}{9}
\bibitem{bikeevo}
\textit{Bicycle Evolution}. November 2007. \\
Online: \url{https://commons.wikimedia.org/wiki/File:Bicycle_evolution-en.svg}

\bibitem{whipple}
F. Whipple. \textit{Stability of the Motion of a Bicycle}. The Quarterly Journal of Pure and Applied Mathematics, Vol XXX, 1899.

\bibitem{papadopoulos}
A. L. Schwab, J. P. Meijaard, J. M. Papadopoulos. \textit{A Multibody Dynamics Benchmark On The Equations of Motion of an Uncontrolled Bicycle}. 5th Nonlinear Dynamics Conferencec, 2005. 

\bibitem{astrom}
K. \AA str{\"o}m, R. Klein, A. Lennartsson. \textit{Bicycle Dynamics and Control: Adapted bicycles for education and research}. IEEE Control Systems, 10.1109/MCS.2005.1499389, September 2005.

\bibitem{fourthorder}
J. M. Papadopoulos, R. S. Hand, A. Ruina. \textit{Bicycle and Motorcycle Balance and Steer Dynamics}. Draft, July 1990.

\bibitem{hinftheory}
D. McFarlane, K. Glover. \textit{Robust Controller Design using Normalised Coprime Factor Plant Descriptions}. Springer Verlag, Lecture Notes in Control and Information Sciences, Vol. 138, 1989.

\bibitem{hinf} 
D. McFarlane, K. Glover. \textit{A Loop Shaping Design Procedure Using $H_{\infty}$ Synthesis}. IEEE Transactions on Automatic Control, Vol. 37 No. 6, June 1992.

\bibitem{nugap}
G. Vinnicombe. \textit{Uncertainty And Feedback: $H_{\infty}$ Loop-Shaping And The $\nu$-Gap Metric}. Imperial College Press, November 2000.

\bibitem{virtualmodel}
J. Pratt, C. Chew, A. Torres, P. Dilworth, G. Pratt. \textit{Virtual Model Control: An Intuitive Approach for Bipedal Locomotion}. International Journal of Robotics Research, Vol. 20 Issue 2, February 2001.

\bibitem{puttingenergy}
R. Ortega, A. van der Schaft, I. Mareels, B. Maschke. \textit{Putting Energy Back in Control}. IEEE Control Systems Magazine, Vol. 21 Issue 2, 2001.

\bibitem{antiwindup}
D. Wilson. \textit{Teaching Your PI Controller to Behave}. April 2013. Online: \\\url{https://e2e.ti.com/blogs_/b/motordrivecontrol/archive/2013/04/13/teaching-your-pi-controller-to-behave-part-vii}

\bibitem{smalluav}
R. W. Beard, T. W. McLain. \textit{Small Unmanned Aircraft: Theory and Practice}. Princeton University Press, 2012.

\end{thebibliography}